---
title: "Lab 8 - Linear Regressions in R"
author: "Sam Csik"
date: "11/20/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Setup
```{r}
##############################
# Load required packages
##############################

library(tidyverse)

##############################
# Import data files
##############################

salmon <- read_csv("salmon_imports_copy.csv")

##############################
# Simplify data
##############################

salmon_simple <- salmon %>% 
  select(year, salmon_fresh)
```

###Create exploratory scatterplot
```{r}
salmon_scatter <- ggplot(data = salmon_simple, aes(x = year, y = salmon_fresh)) +
  geom_point()

# Overall, using a linear regression to describe the relationship between year and fresh salmon imports makes sense. I do not have concerns about using linear regression to describe this relationship, as it is not obviously non-linear despite some strange data that deviate from the general linear trend exhibited. 
```

###Linear Regression
```{r}
##############################
# Perform linear regression to describe the trend in total volume of fresh salmon imported to the US from 1990 to 2017
##############################

lm_salmon <- lm(salmon_fresh ~ year, data = salmon_simple)

# If we were to write out the model equation with the actual variables and coefficients, we would find: 

# Imports(tons) = -14982940 + 7550(year)
```

In the context of this model, the slope (7550) means that for each year that goes by, we see an increase of 7550 tons of salmon import to the US.

In the context of this model, the y-intercept means that at year "0," salmon imports = -14982940 tons. This means that if we were to extrapolate this model for past values, salmon imports would continue to decline to negative values???

###Model diagnostic plots
```{r}
plot(lm_salmon) # view diagnostic plots in console

par(mfrow = c(2,2)) # plot all 4 diagnositcs plots together
plot(lm_salmon)
```

The residuals appear to be normally distributed (qq-plot looks linear, residuals vs fitted looks like points are relatively evenly spread about dashed line (though there are some strange outliers)).

No concerns about heteroscedasticity because there doesn't appear to be any trend in the spread of the residuals. There are 3 outliers.

###Explore model fit and significance
```{r}
summary(lm_salmon)
```

Year significantly predicts salmon imports.

The R2 value means that 94l8% of the variance in salmon inports is predictable from year (in other words, year accurately predicts 94.8% of the variation in salmon imports to the US).

The proportion of variance that is not explained by year is 5.2%

###Prepare a final graph
```{r}
salmon_final_graph <- ggplot(salmon_simple, aes(x = year, y = salmon_fresh)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") + # se displays 95% CI for the predictions of the linear model
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports \n(1,000 points)")
```


###Making predictions using the model for years 2022, 2024, 2026, 2028
```{r}
##############################
# Create df with new values that you want to make predictions for
##############################

new_years <- data.frame(year = c(2022, 2024, 2026, 2028))
                        
##############################
# Feed new_years df into predict
##############################

future_predict <- predict(lm_salmon, newdata = new_years, interval = "confidence") # setting interval = "confidence" reports the upper and lower bounds of the 95% CI for model fit at each year

##############################
# Bind the prediction outcomes (those that are in the "fit" column with the 'new_years' data) to create a more useful table of predicted imports and upper and lower CI
##############################

predictions <- data.frame(new_years, future_predict)
predictions
```

###Find Pearson's r (correlation)
```{r}
##############################
# Use cor.test() function to find Pearson's r for the linear relationship between year and salmon imports
##############################

salmon_r <- cor.test(salmon_simple$year, salmon_simple$salmon_fresh)

# Pearson's r = 0.974
# I would consider this to be a strong positive correlation
```

Year significantly predicts salmon imports (b = 7.5, t(26) = 21.7, *p* < 0.001) with a strong positive correlation between the two (Pearson's *r* = 0.97). The overall model (*salmon imports* = 7.5(*year*) -1.5e7; salmon imports measured in tons) explains a significant amount of variance in salmon imports (F(1,26) = 471.9, *p* < 0.001, *R^2^* = 0.95).
